Models including an interaction of treatment with baseline risk were better able to predict absolute benefit in simulations

### 2021-07-12 submission/manuscript.rmd
use cross-validated performance metrics

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```

@@

constantModel <- fitModelBasedHte(
  data     = gusto, 
  settings = createModelBasedSettings(type = "treatment")
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictBenefitModelBasedHte(
      p             = plogis(.$riskLinearPredictor), 
      modelBasedFit = constantModel
    )
  )

constantC   <- calculateCForBenefit(tmp)
constantIci <- calculateCalibrationForBenefit(tmp)

stratified <- fitStratifiedHte(
  data     = gusto,
  settings = createStratifiedSettings()
)

linearModel <- fitModelBasedHte(
  data     = gusto,
  settings = createModelBasedSettings()
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictBenefitModelBasedHte(
      p             = plogis(.$riskLinearPredictor), 
      modelBasedFit = linearModel
    )
  )

linearC   <- calculateCForBenefit(tmp)
linearIci <- calculateCalibrationForBenefit(tmp)

rcs3Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings()
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(riskLinearPredictor), 
      smoothFit  = rcs3Model
    )
  )

rcs3C   <- calculateCForBenefit(tmp)
rcs3Ici <- calculateCalibrationForBenefit(tmp)

rcs4Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings(nKnots = 4)
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(riskLinearPredictor), 
      smoothFit  = rcs4Model
    )
  )

rcs4C   <- calculateCForBenefit(tmp)
rcs4Ici <- calculateCalibrationForBenefit(tmp)

rcs5Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings(nKnots = 5)
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(.$riskLinearPredictor), 
      smoothFit  = rcs5Model
    )
  )

rcs5C   <- calculateCForBenefit(tmp)
rcs5Ici <- calculateCalibrationForBenefit(tmp)

gustoPerformance <- tibble(
  discrimination = c(
    constantC, linearC, rcs3C,
    rcs4C, rcs5C
  ),
  calibration = c(
    constantIci$ici, linearIci$ici, rcs3Ici$ici,
    rcs4Ici$ici, rcs5Ici$ici
  ),
  method = c(
    "constant treatment effect",
    "linear interaction",
    "RCS smoothing with 3 knots",
    "RCS smoothing with 4 knots",
    "RCS smoothing with 5 knots"
  )
)

cBenefit <- gustoPerformance %>% arrange(discrimination)

@@

We used the risk linear predictor to fit the proposed methods under study for
individualizing absolute benefit predictions. All methods predicted increasing
benefits for patients with higher baseline risk predictions, but the fitted
patterns were clearly different. The adaptive approach selected the model with
RCS smoothing with 4 knots. However, for very low baseline risk this model
predicted decreasing benefit with increasing risk maybe somewhat too flexible.
The more robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice.

>>>>

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
set.seed(19910930)
```

@@

folds <- caret::createFolds(gusto$outcome, k = 4)

evaluatePerformance <- function(data, testRows, settings) {
  isStratified <- FALSE
  if ("modelBased" %in% class(settings)) {
    model <- fitModelBasedHte(data %>% dplyr::slice(-testRows), settings)
    tmp <- data %>%
      slice(testRows) %>%
      dplyr::mutate(
        predictedBenefit = predictBenefitModelBasedHte(
          p             = plogis(.$riskLinearPredictor), 
          modelBasedFit = model
        )
      )

  } else if ("rcs" %in% class(settings)) {
    model <- fitRcsHte(data %>% dplyr::slice(-testRows), settings)
    tmp <- gusto %>%
      slice(testRows) %>%
      dplyr::mutate(
        predictedBenefit = predictSmoothBenefit(
          p          = plogis(riskLinearPredictor), 
          smoothFit  = model
        )
      )
  } else if ("stratified" %in% class(settings)) {
    isStratified <- TRUE
    model <- fitStratifiedHte(data %>% dplyr::slice(-testRows), settings)
    tmp <- gusto %>%
      slice(testRows) %>%
      dplyr::mutate(
        predictedBenefit = predictStratifiedBenefit(
          p          = plogis(riskLinearPredictor), 
          stratifiedHte  = model
        )
      )
  }
  cForBenefit <- calculateCForBenefit(tmp)
  iciForBenefit <- calculateCalibrationForBenefit(tmp)
  aic <- ifelse(
    isStratified,
    NA,
    model$aic
  )

  return(
    list(
      c = cForBenefit,
      ici = iciForBenefit$ici,
      aic = aic
    )
  )
}

constant <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createModelBasedSettings(type = "treatment")
)

constantData <- do.call(bind_rows, constant)
cConstant <- mean(constantData$c)
iciConstant <- mean(constantData$ici)
aicConstnat <- mean(constantData$aic)

stratified <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createStratifiedSettings()
)

stratifiedData <- do.call(bind_rows, stratified)
cStratified <- mean(stratifiedData$c)
iciStratified <- mean(stratifiedData$ici)

linear <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createModelBasedSettings()
)

linearData <- do.call(bind_rows, linear)
clinear <- mean(linearData$c)
iciLinear <- mean(linearData$ici)
aicLinear <- mean(linearData$aic)

rcs3 <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createRcsSettings(nKnots = 3)
)

rcs3Data <- do.call(bind_rows, rcs3)
cRcs3    <- mean(rcs3Data$c)
iciRcs3  <- mean(rcs3Data$ici)
aicRcs3  <- mean(rcs3Data$aic)

rcs4 <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createRcsSettings(nKnots = 4)
)

rcs4Data <- do.call(bind_rows, rcs4)
cRcs4    <- mean(rcs4Data$c)
iciRcs4  <- mean(rcs4Data$ici)
aicRcs4  <- mean(rcs4Data$aic)

rcs5 <- lapply(
  folds,
  evaluatePerformance,
  data = gusto,
  settings = createRcsSettings(nKnots = 5)
)

rcs5Data <- do.call(bind_rows, rcs5)
cRcs5    <- mean(rcs5Data$c)
iciRcs5  <- mean(rcs5Data$ici)
aicRcs5  <- mean(rcs5Data$aic)

gustoPerformance <- tibble(
  discrimination = c(
    cConstant, clinear, cRcs3,
    cRcs4, cRcs5
  ),
  calibration = c(
    iciConstant, iciLinear, iciRcs3,
    iciRcs4, iciRcs5
  ),
  aic = c(
    aicConstnat, aicLinear, aicRcs3,
    aicRcs4, aicRcs5
  ),
  method = c(
    "constant treatment effect",
    "linear interaction",
    "RCS smoothing with 3 knots",
    "RCS smoothing with 4 knots",
    "RCS smoothing with 5 knots"
  )
)

discriminationPerformance <- gustoPerformance %>%
  slice(c(2, 3)) %>%
  select(discrimination)

calibrationPerformance <- gustoPerformance %>%
  slice(c(2, 3)) %>%
  select(calibration)

@@

We used the risk linear predictor to fit the proposed methods under study for
individualizing absolute benefit predictions. All methods predicted increasing
benefits for patients with higher baseline risk predictions, but the fitted
patterns were clearly different. The adaptive approach selected the model with
RCS smoothing with 4 knots. However, for very low baseline risk this model
predicted decreasing benefit with increasing risk may be somewhat too flexible.
The more robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The model with RCS smoothing (3 knots)
had somewhat lower AIC compared to the linear interaction model, slightly better
cross-validated discrimination (c-for-benefit
`r round(discriminationPerformance[2, ], 3)` vs
`r round(discriminationPerformance[1, ], 3)`) and quite similar cross-validated
calibration (ICI-for benefit `r round(calibrationPerformance[2, ], 4)` vs
`r round(calibrationPerformance[1, ], 4)`).

### 2021-07-12 submission/manuscript.tex
use cross-validated performance metrics

We used the risk linear predictor to fit the the proposed methods under
study for individualizing absolute benefit predictions. All methods had
quite comparable results, in the sense that we predicted increasing
benefits for patients with higher baseline risk predictions. In terms of
c-for-benefit (validated internally) all models had quite comparable
performance ranging from 0.519 (RCS smoothing with 3 knots) to 0.542
(RCS smoothing with 4 knots). Similar conclusions could be drawn in
terms of ICI-for-benefit which ranged from 0.0039 (linear interaction
approach) to 0.0053 (RCS smoothing with 4 knots).

The adaptive approach picked the model with RCS smoothing with 4 knots,
which had quite comparable performance to the smooth fit with 5 knots.
However, for lower baseline risk these 2 models predicted implausible
benefits and maybe should be avoided when applying such models in
practice. The linear interaction model and the model with RCS smoothing
(3 knots) made very similar predictions and also followed quite closely
the evolution of the stratified estimates. In this case, we observed
that even using a simple model with a constant relative treatment effect
can result in very similar benefit predictions as other more complex
approaches.

>>>>

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4 knots.
However, for very low baseline risk this model predicted decreasing
benefit with increasing risk may be somewhat too flexible. The more
robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The model with RCS smoothing (3
knots) had somewhat lower AIC compared to the linear interaction model,
slightly better cross-validated discrimination (c-for-benefit 0.537 vs
0.527) and quite similar cross-validated calibration (ICI-for benefit
0.0114 vs 0.0116).
